---
title: "Reproducible Research: Peer Assessment 1"
author: "Ferran Brianso (ferran.brianso@vhir.org)"
date: "13/08/2015"
output: 
  html_document:
    keep_md: true
---

## General settings and library loadings

```{r}
echo = TRUE
library(ggplot2)
```

## Loading and preprocessing the data

Unzip and load csv file
```{r}
unzip("activity.zip")
data <- read.csv("activity.csv", header=TRUE)
```

Preprocess some fields and review the data
```{r}
data$interval <- as.factor(data$interval)
data$date <- as.Date(data$date)
head(data)
tail(data)
dim(data)
```

Subset those cases without NA
```{r}
data.woNA <- na.omit(data)
rownames(data.woNA) <- 1:nrow(data.woNA)
head(data.woNA)
tail(data.woNA)
dim(data.woNA)
```

## What is mean total number of steps taken per day?

Using data set without NAs for this part of the assignment.

Calculate the total number of steps taken per day
```{r}
steps.per.day <- aggregate(data.woNA$steps, list(Date=data.woNA$date), FUN="sum")
head(steps.per.day)
tail(steps.per.day)
```

Make a histogram of the total number of steps taken each day
```{r}
hist(steps.per.day$x, breaks=(length(steps.per.day$x)/3), col="orange",
     main="Histogram of steps per day", xlab="Steps per day")
```

Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(steps.per.day$x)
median(steps.per.day$x)
```

## What is the average daily activity pattern?

Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r}
avg.steps <- aggregate(data.woNA$steps, list(interval=as.numeric(data.woNA$interval)), FUN="mean")
names(avg.steps)[2] <- "avgSteps"
ggplot(avg.steps, aes(interval, avgSteps)) + geom_line(color="orange") + labs(title="Time series plot of the average steps taken per 5-minute interval", x="Intervals", y="Average Number of Steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
avg.steps[which(avg.steps$avgSteps==max(avg.steps$avgSteps)), ]
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(data))
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.  

In this case, the mean for the corresponding 5-minute interval (already stored in avg.steps) has been used to fill each NA steps value.
```{r}
data.filled <- data
head(data.filled)
head(avg.steps)
for (i in 1:nrow(data.filled)) {
      if (is.na(data.filled$steps[i])) {
            interv <- as.numeric(data.filled$interval[i])
            data.filled$steps[i] <- avg.steps[which(avg.steps$interval == interv), ]$avgSteps
      }
}
head(data.filled)
sum(is.na(data.filled))
```

Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


## Are there differences in activity patterns between weekdays and weekends?
